---
title: "Analysis of the Bechdel test and awards on IMBD scores"
author: "William Bevidas, Olachi Mbakwe, Claudia Silverstein"
date: "`r Sys.Date()`"
output: pdf_document
geometry: left=1in,right=1in,top=1in,bottom=1in
urlcolor: blue
header-includes:
  - \usepackage{subfig}
---


The movie industry has been around for over 100 years and along with it has come
rating systems. There are many different types of ratings that can help define
how good a movie is such as Rotten Tomatoes, Letterboxd, IMBD, etc. One category
that does a good job at combining ratings from around the world are IMBD scores. 
Since these are scores from people around the world, we can point out what movies
do better or worse and might even be able to explain some bias in these votes.
One of the ways to measure a movies diversity is through the Bechdel test. There
is also potential for bias in movies that have won awards at award shows.  

We are investing a way to determine bias in IMBD scores through the use of the 
Bechdel test and award winning movies. We have multiple research questions that
we would like to answer:

- Do Bechdel test scores impact IMBD scores?
- Do movie awards impact IMBD scores?
- Does the interaction of Bechdel test scores and number of awards impact the
IMBD score?

We will first discuss some additional background information and describe the
methods we used to collect our data. We will then describe the collected data and
share the results of analysis. Finally, we will close with discussion of results and some post-hoc analysis questions if there is a significant impact such as:

- Does failing the Bechdel test lead to a higher IMBD score?
- Does winning more awards lead to a higher IMBD score?

# Background

## IMBD Scores

The way IMBD scores are calculated is by taking the votes from registered IMBD
users who voted on specific titles in the IMBD database using their account. 
Each person gets one vote per movie. Instead of taking the arithmetic mean,
they use a weighted average based on the votes given and the rating can 
sometimes be influenced on unusual voting activity. Their method for calculating
the final score is not disclosed but they claim that it works the best in
keeping the scores reliable. These scores are helpful in helping people to
decide what to watch, meaning it is a good measure of how good a movie is. 

## Award Ceremonies

Another way of understanding how well a movie can be is through award
ceremonies. Three common award ceremonies that take place yearly are the Oscars,
the Golden Globes, and the Bafta Awards. Getting an award from these ceremonies
shows that you must have done something that was impressive in film, however
this might not always be the case. Award ceremonies and Hollywood have been
known to show bias against minority groups. One of the groups involved has been
women and a big issue is the lack of actresses in the cast. 

In a 2012 study, it was proven that from 1950 to 2006, the number of male main
characters has outnumbered female characters by more than two to one in the most
popular films during that time period. So, more popular movies who will probably
be getting Oscars typically has less females than males. Along with that, a
majority of Oscar voters are males, giving into even more potential bias as a
man might vote for a movie that has more men in it. While there are awards for
best lead actress, we are only going to be interested in awards given to the
entire film, to try and find the underlying bias. 

## The Bechdel Test

The Bechdel test was created by cartoonist Alison Bechdel from one of their 1985
cartoons and is a measure of the representation of women in film and other
fiction. It was meant to call attention to gender inequality in fiction (not
just movies). The rules to pass a test are simple:

- The movie has to have at least two women in it
- These women must talk to each other
- These women must talk about something other than a man

It doesn't seem too hard to pass the film, yet there are many movies out there
that fail. This test is able to break open movies and point out where some
underlying bias is that might not have been noticed before. There can however,
also be skewed results within test. For instance, *The Lord of the Rings: The Two Towers* 
actually passes the test, but only because there is a 5 second clip of a mother
talking to her daughter. Yes there were two women talking and there is some
debate on whether or not it was about a man (the daughter asks where her father
is), but it doesn't mean the film deserved to pass as these two characters are
not relevant to the story. Since this test might not be extremely accurate, it
might be hard to justify any trends but we think it is an easy and useful way 
to recognzie where the underlying bias might be in the movies. 

```{r  Loading data}

bechdel_awards <- read_csv("bechdel_awards.csv")

### Set year to an ordered factor
bechdel_awards$movies.code <- factor(
x = bechdel_awards$movies.code,
levels = c("FAIL", "PASS")
)

### Set awards to an ordered factor with specified levels
bechdel_awards$awards_factor <- cut(
  x = bechdel_awards$awards,
  breaks = c(-Inf, 1, 5, 10, 15, 20),
  labels = c("No Award", "1-5 Awards", "6-10 Awards", "11-15 Awards", "16-20 Awards"),
  include.lowest = TRUE,
  right = FALSE
)

```

## Previous Studies

While there are no studies to our knowledge doing exactly what we are doing,
some people have used the Bechdel test in studies. In a 2021 paper written by
Nelli Abdullaeva titled *The issues of an oversimplified analysis in assessing the representation of women in cinema*, they discuss how passing the Bechdel 
test doesn't necessarily mean it is a good movie or it gives an adequate
representation of women in film. They compared Bechdel test scores with their 
own personal feelings on how the representation of women and compared the two. 
They found that 15 of the 30 movies they looked at passed both the Bechdel 
test, and the personal evaluation. Two movies failed the test but passed the 
evaluation, three passed but failed the personal evaluation, and seven failed 
both. There were a few controversial calls, but overall this simple test showed
that 22 out of the 30 movies were correctly representative based on the 
author's personal feelings. 

# Data Collection

To continue investigation on the IMBD scores, we will make use of 5 different
data sets that have been collected. The first data set consists of a list of 
5,043 IMBD scores and their equivalent movies. There was other unnecessary data
for this report included here as well. The next three data sets included 10,765
different Oscar awards, 7,991 different Golden Globe awards, and 4,176 Bafta
awards that are either a person or movie and that either won that specific
award or did not. The last data set has 1,794 movies each with a binary 
variable on whether or not it passes the Bechdel test. Depending on what movies
match the Bechdel test, we should have a maximum of 1,794 movies to work with
in this study. 

The first step in making a data set to use is finding out how many awards each 
movie won. For each row of data, if the award is True then it got 1 award and 
if is False it got 0. This counts for all types of awards, as long as it
the entire film getting the award and not an actor or actress. Then for each
awards data set, we aggregated the rows and added up the awards to get how many
awards were given to each specific movie at that award ceremony. From there,
the data was joined into one table, where all cases were kept from each of the
three tables. Any cases of missing data between the three were given a value of
0. This is because if a movie was not on another list, it was assumed that it 
wasn't apart of that award show since each of the original lists contained all 
known information. 

Next, the awards are added up from each of the award ceremonies and this final
list consists of 11,003 different movies/people who received awards along with
how many awards they have won. In order to include all movies that have Bechdel
test scores that we have access to, we join these two data sets together, along
with the IMBD score data set. This leaves us with a data set consisting of 
578 observations that include the movie, the number of awards won, the year it 
came out, its Bechdel test score, and its IMBD score. 

The final item to use for analysis is dealing with the amount of awards. Since
there are between 0 and 15 awards, we are going to use categories to simplify the
levels of this factors. We decided to use none, low, medium, and high to determine
the amount of awards given to the movies. None is defined as 0 awards, low is 
between 1 and 5 awards, medium is between 6 and 10 awards, and high is between
11 and 15 awards. 

# Analytical Methods

To analyze our data and answer our research questions we will use R (version 4.3.0)
and make use of ANOVA methods, in particular a full factorial (two-way) ANOVA model.

## Appropriateness of ANOVA

Our response is a continuous variable because it ca

```{r,echo=FALSE,fig.width=4,fig.height=5,fig.align='center'}
# Hasse Diagram
modelLabels <- c("1 Rating Movies 1", "3 Bechdel Score 2", "3 Awards 2", "9 Bechdel Score × Awards 4", "578 Movies 569")
modelMatrix <- matrix(
  data = c(FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE, FALSE),
  nrow = 5,
  ncol = 5,
  byrow = FALSE
)
hasseDiagram::hasse(
 data = modelMatrix,
 labels = modelLabels
)
```


For this report, we will be using IMBD score as the response with the amount of
awards as the factor at 4 levels and the Bechdel test score as the second factor
at 2 levels. 

## Addressing the Multiple Comparisons Problem

Within each hypothesis test, we will use an Unusualness Thresheold equivalent to 
0.1.

# Exploratory Data Analysis

# Results

## Assumptions

## Omnibus

## Post Hoc Analysis

### Comparing Bechdel Scores Given the Amount of Awards

### Comparing Awards Given the Bechdel Score

# Discussion

## Limitations

## Future Work

There is potential here for this project to handle future work. First, every
year new award shows are presented. 

# References

https://en.wikipedia.org/wiki/Bechdel_test#cite_note-7

https://www.jahonline.org/article/S1054-139X(12)00069-9/fulltext

https://help.imdb.com/article/imdb/track-movies-tv/weighted-average-ratings/GWT2DSBYVT2F25SK?ref_=ttrt_wtavg#

https://help.imdb.com/article/imdb/track-movies-tv/ratings-faq/G67Y87TFYYP6TWAV#

https://blogs.fu-berlin.de/abv-gender-diversity/2021/12/13/the-bechdel-test-and-gender-equality-in-the-film-industry/

# Author Contributions


