---
title: "Fruit  files"
author: "Olachi Mbakwe,William Bevidas"
date: "2023-10-31"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	include=FALSE
)

```



```{r}

# Load packages
packages <- c("tidyverse", "knitr","dplyr", "kableExtra","parameters", "emmeans", "DescTools","dunn.test")



lapply(
  X = packages,
  FUN = library,
  character.only = TRUE
)


# Set options
options(knitr.kable.NA = "")


# Load additional tools
source("https://raw.github.com/neilhatfield/STAT461/master/rScripts/ANOVATools.R")

```

# Introduction and Background
<!-- This is where you'll write the portion of your narrative which frames the context for the SRQ. This is also where you will 1) explicitly state the SRQ and 2) incorporate any additional references (if applicable) -->

There have always been theories throughout history on the physiological cost of reproduction in terms of reduced lifespan^[https://www.nature.com/articles/294580a0.pdf]. Although it has not been determined that there is a significant impact on the longevity of human males on reproduction, two scientists, Linda Partidge and Marion Farquhar decided to test the impact on sexual reproduction of male fruit flies. This is because while some species do not need to have a partner to reproduce, fruit flies do, which is why there were chosen for this study. 

In order to make any statistical discoveries within this study, we would need to see that if a male fruit fly reproduces, their lives are cut significantly shorter than those that do not reproduce within their life. Although we cannot determine anything for humans through this study, we can make some inference on if sexual reproduction impacts the longevity of a living object and this could potentially begin a study on how it affects a human life time. While there might be ethical dilemmas involved here and other things to figure out for that part of the research, fruit flies are a good way to begin to understand the affect of sexual activity on longevity. 

Overall, this experiment is trying to help explore whether or not if the reproduction process for male fruit flies affects their longevity? If there turns out to be a statistically significant impact on the fruit flies longevity, our post hoc question would include the following. 

+ Does the amount of other flies in the area make a statistical difference?
+ Does being in an area with virgin fruit flies have a statistical difference from being in an area with pregnant fruit flies?
#Study Design and Methods
In order to investigate the comparative/relationship research question, scientists designed a quasi-experimental study by synthesizing the various versions of studies fruit flies created during the study. Our population of interest is male fruit flies, specifically Drosophila melanogastes. To collect our data, scientists randomly divided 125 male fruit flies into five groups of 25. Each group underwent different conditions of number and type of companions, but all other aspects were kept identical. The longevity was quantitatively collected by the number of days the fly was alive. The scientists then converted the data onto a Numbers spreadsheet to determine balance. The final data collection is publicly available. Our primary response is the longevity of male fruit flies. This has been operationalized as the total number of days they were alive. Our factor of interest is their sexual activity.Taken together and with our research question, ANOVA methods appear to be appropriate.Thus, we will adopt the following null hypothesis: there is no statistically significant impact of sexual activity on longevity. Our alternative hypothesis is then: there is a statistically significant impact of sexual activity on longevity. We may express these hypothesized models as H0 : yij = µ•• + ϵij HA : yij = µ•• + αi + ϵij where yij represents longevity, µ•• the baseline longevity of fruitl fies, αi the additional longevity due to being in factor level i and ϵij, the residuals for each fruit fly j in factor group i. 

type 1 error is point


```{r  Loading data}

fruitflies <- read_csv("fruitflies.csv")

### Set year to an ordered factor
fruitflies$condition <- factor(
x = fruitflies$condition,
levels = c("Null", "One Pregnant", "One Virgin","Eight Pregnant","Eight Virgin")
)

```


```{r}
modelLabels <- c("1 Reproduction 1", "5 Reproduction Condition 4", "125 (Male Fruit Flies) 120")
modelMatrix <- matrix(
  data = c(FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE),
  nrow = 3,
  ncol = 3,
  byrow = FALSE
)
hasseDiagram::hasse(
 data = modelMatrix,
 labels = modelLabels
)
```

